---
title: 'Cogsci2018-OSM'
author: "Derek Powell, Kara Weisman, Ellen Markman"
date: "February 1, 2018"
output: html_notebook
---

```{r}
# moved all data import/munging to a separate file 
# so we can have multiple workbooks working withthe same data

source("../Study1/vacc_import_data.R")

library(tidyverse)
library(stringr)
library(corrplot)
library(viridis)
library(psych)

## set the seed to make your partition reproductible
set.seed(123)
trainInd <- sample(seq_len(nrow(d_bn)), size = floor(nrow(d_bn)*.80))

train <- d_bn[trainInd, ]
test <- d_bn[-trainInd, ]
```

# Introduction

This is a notebook for the first experiment of our (Kara, Ellen, and Derek) study examining vaccine theories using multiple attitude measures. We recruited 1200 and are left with `r length(d_demo$ResponseId)` after excluding any participants who failed at least 1 of the 2 attention checks.


# Demographics

A quick look at demographics variables...

## Study duration

```{r duration}
ggplot(d_demo, aes(x = Duration/60)) +
  geom_histogram() +
  geom_vline(xintercept = median(d_demo$Duration/60), color = "blue") +
  annotate("label", x = 0, y = 350, color = "blue", hjust = 0,
           label = paste0("median: ", round(median(d_demo$Duration/60)), "min")) +
  labs(x = "duration (min)") +
  theme_bw()
```

## Gender

```{r sex}
ggplot(d_demo, aes(x = sex)) +
  geom_bar() +
  labs(x = "gender") +
  theme_bw() +
  coord_flip()
```

## Age

```{r age}
ggplot(d_demo, aes(x = age)) +
  geom_histogram() +
  geom_vline(xintercept = median(d_demo$age), color = "blue") +
  annotate("label", x = 0, y = 150, color = "blue", hjust = 0,
           label = paste0("median: ", round(median(d_demo$age)), "y")) +
  labs(x = "age (y)") +
  theme_bw()
```

## Race

```{r race}
ggplot(d_demo, aes(x = race)) +
  geom_bar() +
  labs(x = "race") +
  theme_bw() +
  coord_flip()
```

## Religion

```{r religion}
ggplot(d_demo, aes(x = religion)) +
  geom_bar() +
  labs(x = "religion") +
  theme_bw() +
  coord_flip()
```

## Education

```{r education}
ggplot(d_demo, aes(x = educ)) +
  geom_bar() +
  labs(x = "educ") +
  theme_bw() +
  coord_flip()
```

## Income

```{r income}
ggplot(d_demo, aes(x = income)) +
  geom_bar() +
  labs(x = "income") +
  theme_bw() +
  coord_flip()
```

## Parent

```{r parent}
ggplot(d_demo, aes(x = parent)) +
  geom_bar() +
  labs(x = "parent") +
  theme_bw() +
  coord_flip()
```

## Expecting

```{r expecting}
ggplot(d_demo, aes(x = expecting)) +
  geom_bar() +
  labs(x = "expecting") +
  theme_bw() +
  coord_flip()
```

## Number of children

```{r children}
ggplot(d_demo, aes(x = factor(children))) +
  geom_bar() +
  labs(x = "number of children") +
  theme_bw() +
  coord_flip()
```

## Age of youngest child

```{r youngest_child}
ggplot(d_demo, aes(x = factor(youngest_child))) +
  geom_bar() +
  labs(x = "age of youngest child") +
  theme_bw() +
  coord_flip()
```

## Comments on study experience

```{r comments}
d_demo %>% select(comments)
```

# Scale reliability analyses

As a reminder, our goal was $\alpha$ >= 0.80 for each scale individually.

## Alpha for each scale

### Disease rarity (*diseaseRare*)

```{r reliability diseaseRare}
alpha_diseaseRare <- d_long %>%
  filter(question_block == "diseaseRare") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_diseaseRare
```

For *diseaseRare*, $\alpha$ = `r round(alpha_diseaseRare$total$raw_alpha, 2)`. This is *just* under our cutoff (ok).

### Disease severity (*diseaseSevere*)

```{r reliability diseaseSevere}
alpha_diseaseSevere <- d_long %>%
  filter(question_block == "diseaseSevere") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_diseaseSevere
```

For *diseaseSevere*, $\alpha$ = `r round(alpha_diseaseSevere$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Holistic balance (*hb*)

```{r reliability hb}
alpha_hb <- d_long %>%
  filter(question_block == "hb") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_hb
```

For *hb*, $\alpha$ = `r round(alpha_hb$total$raw_alpha, 2)`. This is a little over our cutoff (good).

### Infant immune system has limited capacity (*infantImmLimCap*)

```{r reliability infantImmLimCap}
alpha_infantImmLimCap <- d_long %>%
  filter(question_block == "infantImmLimCap") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_infantImmLimCap
```

For *infantImmLimCap*, $\alpha$ = `r round(alpha_infantImmLimCap$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Infant immune system is weak (*infantImmWeak*)

```{r reliability infantImmWeak}
alpha_infantImmWeak <- d_long %>%
  filter(question_block == "infantImmWeak") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_infantImmWeak
```

For *infantImmWeak*, $\alpha$ = `r round(alpha_infantImmWeak$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Medical skepticism (*medSkept*)

```{r reliability medSkept}
alpha_medSkept <- d_long %>%
  filter(question_block == "medSkept") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_medSkept
```

For *medSkept*, $\alpha$ = `r round(alpha_medSkept$total$raw_alpha, 2)`. This is just under our cutoff (ok).

### Naturalism (*nat*)

```{r reliability nat}
alpha_nat <- d_long %>%
  filter(question_block == "nat") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_nat
```

For *nat*, $\alpha$ = `r round(alpha_nat$total$raw_alpha, 2)`. This is just under our cutoff (ok).

### Overparenting (*overpar*)

```{r reliability overpar}
alpha_overpar <- d_long %>%
  filter(question_block == "overpar") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_overpar
```

For *overpar*, $\alpha$ = `r round(alpha_overpar$total$raw_alpha, 2)`. This is quite a bit under our cutoff (not great).

### Parental expertise (*parentExpert*)

```{r reliability parentExpert}
alpha_parentExpert <- d_long %>%
  filter(question_block == "parentExpert") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_parentExpert
```

For *parentExpert*, $\alpha$ = `r round(alpha_parentExpert$total$raw_alpha, 2)`. This is just over our cutoff (good).

### Vaccines are dangerous (*vaccDanger*)

```{r reliability vaccDanger}
alpha_vaccDanger <- d_long %>%
  filter(question_block == "vaccDanger") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccDanger
```

For *vaccDanger*, $\alpha$ = `r round(alpha_vaccDanger$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Vaccines are effective (*vaccEff*)

```{r reliability vaccEff}
alpha_vaccEff <- d_long %>%
  filter(question_block == "vaccEff") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccEff
```

For *vaccEff*, $\alpha$ = `r round(alpha_vaccEff$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Vaccine intentions (*vaccIntent*)

```{r reliability vaccIntent}
alpha_vaccIntent <- d_long %>%
  filter(question_block == "vaccIntent") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccIntent
```

For *vaccIntent*, $\alpha$ = `r round(alpha_vaccIntent$total$raw_alpha, 2)`. This is just over our cutoff (good).

### Vaccines strain the immune system (*vaccStrain*)

```{r reliability vaccStrain}
alpha_vaccStrain <- d_long %>%
  filter(question_block == "vaccStrain") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccStrain
```

For *vaccStrain*, $\alpha$ = `r round(alpha_vaccStrain$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Vaccines are toxic (*vaccTox*)

```{r reliability vaccTox}
alpha_vaccTox <- d_long %>%
  filter(question_block == "vaccTox") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccTox
```

For *vaccTox*, $\alpha$ = `r round(alpha_vaccTox$total$raw_alpha, 2)`. This is well over our cutoff (good).

## Scale reliability summary

Things worked out pretty well:

- 10 scales were above the $\alpha$ = 0.80 cutoff (*diseaseSevere*, *hb*,*infantImmLimCap*, *infantImmWeak*, *parentExpert*, *vaccDanger*, *vaccEff*, *vaccIntent*, *vaccStrain*, *vaccTox*)
- 3 scales were just under the cutoff (*diseaseRare*: $\alpha$ = `r round(alpha_diseaseRare$total$raw_alpha, 2)`; *medSkept*: $\alpha$ = `r round(alpha_medSkept$total$raw_alpha, 2)`; *nat*: $\alpha$ = `r round(alpha_nat$total$raw_alpha, 2)`)
- 1 scale was pretty well below the cutoff (*overpar*: $\alpha$ = `r round(alpha_overpar$total$raw_alpha, 2)`)

# Scale score distributions

Let's look at the distributions of scores on each scale. I scored each scale by taking the average of the participant's responses to items on that scale - so for each scale, scores could theoretically range from -3 (participant said "Strongly disagree" to all items after reverse-coding) to +3 (participant said "Strongly agree" to all items after reverse-coding).

```{r score distributions}
score_medians <- d_sum %>%
  group_by(question_block) %>%
  summarise(median_score = median(score))

ggplot(d_sum, aes(x = score)) +
  facet_wrap(~ question_block, ncol = 5) +
  geom_histogram(bins = 25) +
  # scale_x_continuous(limits = c(-3, 3), breaks = seq(-3, 3, 1)) +
  geom_vline(data = score_medians, aes(xintercept = median_score), color = "blue") +
  geom_label(data = score_medians,
             aes(x = -3, y = 200, label = paste0("median: ", round(median_score, 2))),
             color = "blue", hjust = 0) +
  labs(x = "score (min: -3, max: 3)") +
  theme_bw()
```

Most of these scales are pretty normally distributed, with the major exceptions of *diseaseSevere* and *vaccIntent*, which are both severely skewed, and a few other more minor ceiling/floor issues (e.g., *vaccDanger*, *vaccEff*, *vaccTox*).


